Curve counting in genus one: elliptic singularities {\it \&} relative geometry

### 2019-06-28 EllipticSingularitiesSplitting.tex
finished my description of D from Dtilde

\begin{proof}
Choose a component $C^\prime$ of the curve which lies on the circle. Then $\OO(\delta)$ is equal to the tensor product of tangent line bundles along the path of edges connecting $C^\prime$ to the core. However, this tensor product of line bundles is telescoping: any vertex which appears in the middle of the path will have an incoming and outgoing node, and since the vertex corresponds to a smooth rational curve these tangent spaces are naturally dual to each other~\cite[\S 2.2]{VZ}. Thus these factors cancel, and the result follows.
%Note that the final term on the right hand side is not given by $\TT_{q_i} C_0$ (where $q_i$ is the appropriate splitting node) but rather the tangent space to the core. 
\end{proof}
The content of Lemma \ref{lemma fibres of Odelta} is that the choice of point in the fibre of $\widetilde\Dcal \to \widetilde\Ecal$ provides a natural identification between the inward-pointing tangent spaces on the circle. This identification amounts to a specific linear dependence relation between the tangent vectors, which gives a point in the moduli space of attaching data for the corresponding elliptic singularity \cite[\S 2.2]{SMY2}. We may therefore express the factorisation condition in terms of the following natural morphism of vector bundles on $\widetilde\Dcal$:
\begin{equation*} \Sigma \operatorname{d}\!f \colon \OO(\delta)\otimes TC_0^\vee \to \ev_q^\st T H. \end{equation*}
Here $\Sigma \operatorname{d}\!f$ is defined by summing the images of the tangent vectors at the inward-pointing nodes on the circle, using the natural identification of each of these tangent spaces with $\OO(\delta)\otimes TC_0^\vee$ established above. This gives a section of the vector bundle
\begin{equation*} V = (\ev_q^\st TH) \otimes (TC_0) \otimes \OO(-\delta) \end{equation*}
which is transverse and whose vanishing locus coincides with $\Dcal$. Thus we obtain:
\begin{prop} \label{class of D} $[\Dcal] = \operatorname{e}(V) \cap [\widetilde\Dcal]$.\end{prop}
We claim that the class $\operatorname{e}(V)$ is tautological and computable. The only part which needs justification is $\OO(\delta)$. We claim that this can be obtained by pulling back a line bundle from $\Ecal$ and twisting by exceptional divisors.
\begin{lemma} For $i\in \{1,\ldots,r\}$ consider the line bundle
\begin{equation*} T_i = T_{q_i} C_i \otimes T_{q_i} C_0 \end{equation*}
on $\Ecal$, where $q_i$ is the splitting node. Then $\OO(\delta)$ is isomorphic to the pull-back of $T_i$ to $\widetilde\Dcal$ twisted by all exceptional divisors.\end{lemma}
We illustrate this lemma in the following example, which brings out the interplay between logarithmic moduli and blow-ups.
\begin{example}
Consider the following ray $\tau$ and sublocus $\rho$:
\begin{figure}[h]

\begin{minipage}{0.3\textwidth}
\centering
\begin{tikzpicture}
\draw [red] plot [smooth cycle] coordinates {(0,-2) (-0.5,-1) (0,0) (1,0.5) (3,-1) (1,-2.5)};

@@

\draw[color=blue,->] (0,-3.5) -- (3,-3.5);
\draw[fill=blue] (0,-3.5) circle[radius=2pt];
\draw[fill=blue] (2,-3.5) circle[radius=2pt];
\end{tikzpicture}
\caption{$\tau$}
\end{minipage}\hfill
\begin{minipage}{0.7\textwidth}
\centering
\begin{tikzpicture}
\draw [red] plot [smooth cycle] coordinates {(-4,-2) (-2,-1) (-2,1) (0,1) (3,-1) (-1,-3.5)};

@@

\noindent On the cone $\rho$ we have the following tropical equations
\begin{align*}
e:=e_1 = e_2 \\
f:=f_1 = f_2 = 2f_3 \\
g:=g_1 = g_2 = 2g_3
\end{align*}
together with the equation imposed by the choice of alignment:
\begin{align*} & e_1 + f_1 = e_2 + f_3 + g_3 \\
\Leftrightarrow \ & e + f = e + f/2 + g/2 \\
\Leftrightarrow \ & f = g.
\end{align*}
Notice that setting $f=g=0$ we recover $\tau$. We claim that if we twist the bundle
\begin{equation*} E = \left( \TT_{q_1} C_1 \otimes \TT_{q_1} C_0 \right) \oplus \left( \TT_{q_2} C_2 \otimes \TT_{q_2} C_0 \right) \end{equation*}
by the line bundle corresponding to the divisor in $\widetilde\Dcal$ defined by $\rho$, we will obtain the bundle
\begin{equation*} F = \left( \TT_{f_1}C_1^1 \otimes \TT_{e_1} C_0 \right) \oplus \left( \TT_{g_2} C_2^2 \otimes \TT_{e_2} C_0 \right). \end{equation*}
described earlier. The line bundle corresponding to $\rho$ is simply $\OO(f)=\OO(g)$. We have:
\begin{align*} \left( \TT_{q_1} C_1 \otimes \TT_{q_1} C_0 \right) \otimes \OO(f) & = \TT_{q_1}C_1 \otimes \TT_{q_1} C_0 \otimes \TT_{f_1} C_1^0 \otimes \TT_{f_1} C_1^1 \\
& = \TT_{e_1} C_1^0 \otimes \TT_{e_1} C_0 \otimes \TT_{f_1} C_1^0 \otimes \TT_{f_1} C_1^1\\
&  = \TT_{f_1} C_1^1 \otimes \TT_{e_1} C_0\end{align*}
where we have used the telescoping trick $\TT_{f_1}C_1^0 \otimes \TT_{e_1} C_1^0 \cong \OO$. On the other hand, for the $C_2$ summand we write $f=f/2 + g/2=f_3+g_3$ to obtain:
\begin{align*} \left( \TT_{q_2}C_2 \otimes \TT_{q_2}C_0\right) \otimes \OO(f) & = \TT_{e_2} C_0 \otimes \TT_{e_2} C_2^0 \otimes \OO(f_3) \otimes \OO(g_3) \\
& = \TT_{e_2} C_0 \otimes \TT_{e_2} C_2^0 \otimes \TT_{f_3} C_2^0 \otimes \TT_{f_3} C_2^1 \otimes \TT_{g_3} C_2^1 \otimes \TT_{g_3} C_2^2 \\
& = \TT_{g_3} C_2^2 \otimes \TT_{e_2} C_0.\end{align*}
This observation generalises, giving a precise relation between the intrinsically-defined vector bundle $F$ on $\widetilde\Dcal$ (which we use to impose the factorisation condition) and the the vector bundle $E$ pulled back from $\Ecal$. This allows us to express the Chern classes of $F$ in terms of pull-backs of tautological classes from $\Ecal$ and boundary strata on $\widetilde\Dcal$, which is necessary to compute integrals.
\end{example}

>>>>

\begin{proof}
Choose a component $C^\prime$ of the curve which lies on the circle. Then $\OO(\delta)$ is equal to the tensor product of tangent line bundles along the path of edges connecting $C^\prime$ to the core. However, this tensor product of line bundles is telescoping: any vertex which appears in the middle of the path will have an incoming and outgoing node, and since the vertex corresponds to a smooth rational curve these tangent spaces are naturally dual to each other~\cite[\S 2.2]{VZ}. Thus these factors cancel, and the result follows.
%Note that the final term on the right hand side is not given by $\TT_{q_i} C_0$ (where $q_i$ is the appropriate splitting node) but rather the tangent space to the core. 
\end{proof}
The content of Lemma \ref{lemma fibres of Odelta} is that the choice of point in the fibre of $\widetilde\Dcal \to \widetilde\Ecal$ provides a natural identification between the inward-pointing tangent spaces on the circle. This identification amounts to a specific linear dependence relation between the tangent vectors, which gives a point in the moduli space of attaching data for the corresponding elliptic singularity \cite[\S 2.2]{SMY2}. We may therefore express the factorisation condition in terms of the following natural morphism of vector bundles on $\widetilde\Dcal$:
\begin{equation*} \Sigma \operatorname{d}\!f \colon \OO(\delta)\otimes TC_0^\vee \to \ev_q^\st T H. \end{equation*}
Here $\Sigma \operatorname{d}\!f$ is defined by summing the images of the tangent vectors at the inward-pointing nodes on the circle, using the natural identification of each of these tangent spaces with $\OO(\delta)\otimes TC_0^\vee$ established above. This gives a section of the vector bundle
\begin{equation*} V = (\ev_q^\st TH) \otimes (TC_0) \otimes \OO(-\delta) \end{equation*}
which is transverse and whose vanishing locus coincides with $\Dcal$. Thus we obtain:
\begin{prop} \label{class of D} $[\Dcal] = \operatorname{e}(V) \cap [\widetilde\Dcal]$.\end{prop}
We claim that the class $\operatorname{e}(V)$ is tautological and computable. The only part which needs justification is $\OO(\delta)$. We claim that this can be obtained by pulling back a line bundle from $\Ecal$ and twisting by exceptional divisors.
\begin{lemma} For $i\in \{1,\ldots,r\}$ consider the line bundle
\begin{equation*} T_i = T_{q_i} C_i \otimes T_{q_i} C_0 \end{equation*}
on $\Ecal$, where $q_i$ is the splitting node. Then $\OO(\delta)$ is isomorphic to the pull-back of $T_i$ to $\widetilde\Dcal$ twisted by all exceptional divisors.\end{lemma}
We illustrate this lemma in the following example, which brings out the interplay between logarithmic moduli and blow-ups.
\begin{example}
Consider the locus $\Ecal$ given by $\sigma$ below, and the exceptional divisor $\widetilde\Dcal_\rho \subseteq \widetilde\Dcal$ given by $\rho$:
\begin{figure}[!htbp]
\begin{minipage}{0.3\textwidth}
\centering
\begin{tikzpicture}
%\draw [red] plot [smooth cycle] coordinates {(0,-2) (-0.5,-1) (0,0) (1,0.5) (3,-1) (1,-2.5)};

@@

\draw[color=blue,->] (0,-3.5) -- (3,-3.5);
\draw[fill=blue] (0,-3.5) circle[radius=2pt];
\draw[fill=blue] (2,-3.5) circle[radius=2pt];
\end{tikzpicture}
\caption{$\sigma$}
\end{minipage}\hfill
\begin{minipage}{0.7\textwidth}
\centering
\begin{tikzpicture}
\draw [red] plot [smooth cycle] coordinates {(-4,-2) (-2,-1) (-2,1) (0,1) (3,-1) (-1,-3.5)};

@@

On the cone $\rho$ we have the following tropical continuity equations
\begin{align*}
e:=e_1 = e_2, \qquad f:=f_1 = f_2 = 2f_3, \qquad g:=g_1 = g_2 = 2g_3
\end{align*}
together with the equation imposed by the choice of alignment:
\begin{align*} e_1 + f_1 = e_2 + f_3 + g_3 \qquad (\Longleftrightarrow f = g).
\end{align*}
From now on we use the same symbol to denote an edge length and the corresponding node on $\widetilde\Dcal_\rho$. On this locus we have line bundles pulled back from $\Ecal$
\begin{align*} T_1 & = T_{e_1} C_1^0 \otimes T_{e_1} C_0 \\
T_2 & = T_{e_2} C_2^0 \otimes T_{e_2} C_0\end{align*}
as well as the logarithmic line bundle
\begin{equation*} \OO(\delta) = T_{f_1} C_1^1 \otimes T_{e_1} C_0 = T_{f_2} C_1^2 \otimes T_{e_1} C_0 = T_{g_3} C_2^2 \otimes T_{e_2}C_0 
\end{equation*}
where the identifications follow from Lemma \ref{lemma fibres of Odelta}. We claim that: 
\begin{equation} \label{Odelta eqn} \OO(\delta) = T_1\otimes \OO(\widetilde\Dcal_\rho) = T_2 \otimes \OO(\widetilde\Dcal_\rho).\end{equation}
Notice that $\OO(\widetilde\Dcal_\rho) = \OO(f)=\OO(g)$. Thus we have
\begin{align*} T_1 \otimes \OO(\widetilde\Dcal_\rho) & = \OO(f_1) \otimes T_{e_1} C_1^0 \otimes T_{e_1}C_0 \\
& = T_{f_1}C_1^1 \otimes T_{f_1} C_1^0 \otimes T_{e_1} C_1^0 \otimes T_{e_1}C_0  \\
& = T_{f_1}C_1^1\otimes T_{e_1}C_0 \qquad (= \OO(\delta))
\end{align*}
where in the last line we have used the same telescoping trick $\TT_{f_1}C_1^0 \otimes \TT_{e_1} C_1^0 = \OO$ used in the proof of Lemma \ref{lemma fibres of Odelta}. On the other hand, for $T_2$ we may write $f=f/2 + g/2=f_3+g_3$ in order to obtain:
\begin{equation*} T_2 \otimes \OO(\widetilde\Dcal_\rho) = \OO(g_3) \otimes \OO(f_3) \otimes T_{e_2} C_2^0 \otimes T_{e_2} C_0 = T_{g_3}C_2^2 \otimes T_{e_2}C_0 = \OO(\delta) \end{equation*}
where we have used the same telescoping trick. Thus we have proven \eqref{Odelta eqn}. This observation generalises, giving a precise relation between the intrinsically-defined line bundle $\OO(\delta)$ on $\widetilde\Dcal$ (which we use to impose the factorisation condition) and the the line bundles $T_i$ pulled back from $\Ecal$. This allows us to express the Chern classes of $\OO(\delta)$ (and hence $V$) in terms of pull-backs of tautological classes from $\Ecal$ and boundary strata on $\widetilde\Dcal$, which is sufficient in order to compute integrals.
\end{example}
